---
title: Model 3 simulation results
---

## Run simulation

```{r}
source("../scripts/dynamics_model3.r")
consts <- expand.grid(
    n_id = c(1000),
    scenario = 1,
    time_start = 0,
    time_end = 100,
    time_steps = 1,
    parm_n = 2,
    disease_threshold = 0.45
)

cond <- tibble(
    scenario = 1,
    n_phen = 4,
    variables = c("p", "a", "m", "f"),
    phen_mean = c(1,7.5,3,3),
    phen_var = c(1,7.5,3,3),
    phen_h2 = c(0.8, 0.8, 0.8, 0.8),
    genotype_af = c(0.4, 0.4, 0.4, 0.4)
)

consts %>% unlist() %>% sum() %>% set.seed()

starting_conditions <- simulate_starting_conditions(
    n_id=consts$n_id, 
    phen_names=cond$variables, 
    phen_mean=cond$phen_mean, 
    phen_var=cond$phen_var, 
    phen_h2=cond$phen_h2, 
    genotype_af=cond$genotype_af
)

starting_conditions$phen <- t(t(starting_conditions$phen) + abs(apply(starting_conditions$phen,2,min))+apply(starting_conditions$phen,2,max)*0.001)

cond_p <- tibble(
    scenario = 1,
    n_phen = 14,
    variables = c("Pbp", "Ppp", "Pfp", "App", "Aph", "Afp", "Afh", "Mbp", "Mpp", "Mph", "Fdam", "gp", "gm", "gf"),
    par_mean = c(0.25,200,25,75,1.5,80,2.5,0.05,25,5,0.05,5,5,2.5),
    par_var = c(0.25,200,25,75,1.5,80,2.5,0.05,25,5,0.05,5,5,2.5),
    par_h2 = c(0.8, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8),
    genotype_af = c(0.4, 0.4, 0.4, 0.4, 0.4, 0.4, 0.4, 0.4, 0.4, 0.4, 0.4, 0.4, 0.4, 0.4)
)

params <- simulate_params(
    n_id=consts$n_id, 
    par_names=cond_p$variables, 
    par_mean=cond_p$par_mean, 
    par_var=cond_p$par_var, 
    par_h2=cond_p$par_h2, 
    genotype_af=cond_p$genotype_af
)

params$par <- t(t(params$par) + abs(apply(params$par,2,min))+apply(params$par,2,max)*0.001)

dyn <- simulate_dynamics(
    starting_conditions = starting_conditions$phen,
    params = params$par,
    n = consts$parm_n,
    disease_threshold = consts$disease_threshold,
    time_start = consts$time_start,
    time_end = consts$time_end,
    time_steps = consts$time_steps
)

save(dyn, file = paste0("../results/","sim_data_model3.RData"))

```

Example phenotype patterns over time (e.g. individual 1:

```{r}
plot_dynamics(dyn, starting_conditions$phen, 1)
```

## Genetic analysis

There are 4 genotypes, one for each of `m`, `p`, `a` and `f`. 

We can see how each genotype associates with each trait over time

```{r}
gres <- pblapply(unique(dyn$time), function(T)
{
    dyn1 <- subset(dyn, time == T)
    gres <- lapply(c("p", "a", "f", "m"), function(v)
    {
        gwas(dyn1[[v]], starting_conditions$geno) %>%
        mutate(snp=cond$variables, trait=v)
    }) %>% 
        bind_rows() %>%
        mutate(time=T)
    return(gres)
}) %>% bind_rows()

gres$pval[gres$pval < 1e-50] <- 1e-50

ggplot(gres, aes(y=-log10(pval), x=time)) +
geom_line(aes(colour=snp, linetype = snp != trait)) +
facet_grid(paste0("trait = ",trait) ~ .) +
scale_colour_brewer(type="qual", palette=3) +
labs(x="Time", y="SNP-trait assoc, -log10 p-value", colour="SNP", linetype="Pleiotropy")
```

Another view, aggregating the results over time:

```{r}
gres %>% group_by(trait, snp) %>%
    mutate(fdr=p.adjust(pval, "fdr")) %>%
    summarise(n=n(), power=sum(fdr < 0.05)/n) %>%
    mutate(snp=paste0("G", snp)) %>%
    ggplot(., aes(x=snp, y=power)) +
        geom_bar(stat="identity", aes(fill=snp), position=position_dodge(width=1)) +
        scale_fill_brewer(type="qual", palette=3) +
        facet_grid(. ~ trait)
```

Summary:

1. The SNP-trait patterns seem to be very similar across all traits, suggesting that the traits are all highly correlated
2. Genetic associations are highly time-dependent, with power depending on the point of the cycle being measured
3. `f` (Fibronectin) looks like its SNP seldom associates. Is there an obvious reason why?


## MR instruments

Perform MR of each trait against every other trait, and at each timepoint


```{r}
analyses <- expand.grid(
    exposure = c("p", "a", "f", "m"),
    outcome = c("p", "a", "f", "m"),
    time = unique(dyn$time)
) %>% filter(exposure != outcome)

gr <- pblapply(1:nrow(analyses), function(i)
{
    dyn1 <- subset(dyn, time == analyses$time[i])
    get_effs(dyn1[[analyses$exposure[i]]], dyn1[[analyses$outcome[i]]], starting_conditions$geno, analyses$exposure[i], analyses$outcome[i]) %>%
        mutate(
            SNP=cond$variables,
            dir_est = rsq.exposure > rsq.outcome,
            inst = SNP == exposure
        ) %>%
        summarise(
            time = analyses$time[i],
            exposure=exposure[1],
            outcome=outcome[1],
            n_instruments = sum(pval.exposure < 1e-3),
            n_instruments_steiger = sum(pval.exposure < 1e-3 & dir_est),
            true_inst_sig = sum(pval.exposure < 1e-3 & SNP == exposure),
            true_inst_sig_steiger = sum(pval.exposure < 1e-3 & SNP == exposure & dir_est)
        ) %>%
        return()
}) %>% bind_rows()
```

For each exposure, averaging across all MR analyses what types of instruments are being identified?

```{r}
gr %>% tidyr::pivot_longer(
    cols=c(n_instruments, n_instruments_steiger, true_inst_sig, true_inst_sig_steiger)
) %>%
    group_by(exposure, name) %>%
    summarise(value=mean(value)) %>%
    mutate(
    	name=case_when(
    		name == "n_instruments" ~ "Number of discovered instruments",
    		name == "n_instruments_steiger" ~ "Number of discovered instruments after Steiger filtering",
    		name == "true_inst_sig" ~ "True instrument discovered",
    		name == "true_inst_sig_steiger" ~ "True instrument discovered after Steiger filtering"
    	)
    )
```

Summary:

1. Traits often have instruments which are not only their direct genotypic effects (e.g. trait `a` often finds 2 instruments, one of which is its direct effect)
2. Steiger filtering seems to actually make things worse - which suggests that feedback loops alter the directionality of the genotype strengths (e.g. trait `a` tends to drop its direct instrument after Steiger filtering just as often as it drops its pleiotropic effect)

